---
title: "diabetes_analysis"
output: pdf_document
date: "2023-08-26"
---

### The dataset contains information about Pima Indian women, and it is often used to build predictive models to determine whether a person has diabetes based on certain features or risk factors. Carrying my hypotheses over from the Excel document: H0: There is no statistically significant relationship between bmi and diabetes outcome.

### HA: There IS statistically significanct relationship between bmi and diabetes outcome.

Step 1: Imports

Import libraries, import data

```{r}
library(tidyverse)
df <- read.csv("/Users/alexanderbriody/Desktop/Coding Temple/capstone_1/diabetes_analysis_cap1/diabetes.csv")

head(df)
View(df)
```

```{r}
# Cleaning the data by replacing all "." with "_" and coverting to lowercase for column names
clean_fun <- function(x){
  names(x) <- names(x) %>% tolower()
  return(x)
}

df <- clean_fun(df)
View(df)
```

## Step2: Tidy the Data

Check for data types following format

Check and replace null values

```{r}
df %>% summarize(na_count = sum(is.na(df)))
colSums(is.na(df))
```

Step 3: Explore

Gather information:

Decide on independent/dependent variables

Decide on a stat/model test

Visualizations of columns

Descriptive statistics

```{r}
head(df)
```

```{r}
# Outcome is the target variable; 0 for no diabetes, 1 for diabetes.
# Pregnancies: Number of times the woman has been pregnant.
cor(df$outcome, df$pregnancies)
# A low correlation score between these two variables indicates that there is little to no linear relationship between them.
```

```{r}
# Glucose: Plasma glucose concentration in an oral glucose tolerance test.
cor(df$outcome, df$glucose)
# A .4666 correlation is considered a medium correlation. A medium correlation indicates a moderate linear relationship between the variables. This is line with the Excel line graph between glucose concentration and bmi, suggesting a statistical relationship. 
```

```{r}
# BloodPressure: Diastolic blood pressure (mm Hg).
cor(df$outcome, df$bloodpressure)
# A low correlation score between these two variables indicates that there is little to no linear relationship between them.
```

```{r}
# SkinThickness: Triceps skinfold thickness (mm).
cor(df$outcome, df$skinthickness)
# A low correlation score between these two variables indicates that there is little to no linear relationship between them.
```

```{r}
# Insulin: 2-Hour serum insulin (mu U/ml).
cor(df$outcome, df$insulin)
# A low correlation score between these two variables indicates that there is little to no linear relationship between them.
```

```{r}
# BMI: Body mass index (weight in kg / (height in meters)^2).
cor(df$outcome, df$bmi)
# A .3000 correlation is considered at the cusp of a medium correlation. A medium correlation indicates a moderate linear relationship between the variables. 
```

```{r}
# DiabetesPedigreeFunction: A function that scores the likelihood of diabetes based on family history.
cor(df$outcome, df$diabetespedigreefunction)
# A low correlation score between these two variables indicates that there is little to no linear relationship between them.
```

```{r}
# Age: Age in years.
cor(df$outcome, df$age)
# A low correlation score between these two variables indicates that there is little to no linear relationship between them.
```

```{r}
ggplot(df, aes(x = outcome, y = bmi)) +
  geom_jitter(aes(color = factor(outcome)), alpha = 0.5) +
  labs(x = "outcome", y = "bmi") +
  scale_color_discrete(name = "outcome", labels = c("no diabetes", "diabetes")) 

# This graph shows that there is no clear cut relationshp between diabetes outcome and bmi. While some higher bmi's is associated with diabetes outcome, many higher bmi's are also not associated with diabetes, though small portion of the top bmi's are associated with diabetes outcome.
```

```{r}

ggplot(df, aes(x = bmi, y = outcome)) +
  geom_point(aes(color = factor(outcome)), alpha = 0.5) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"),  color = "blue") +
  labs(x = "bmi", y = "outcome") +
  scale_color_discrete(name = "outcome", labels = c("no diabetes", "diabetes")) 

# This graph tells me about the relationship between bmi and diabetes outcome. The best fit curve is a sigmoid between a x-variable (the predictor) and a binary, target, dependent variable - the diabetes outcome. It does visually frame the diabetes outcome to be associated with higher bmi's in general, but there is still a lot of overlap between the two diabetes outcomes for the bmi's.
```

```{r}
summary(glm(df$outcome ~ df$bmi))
# However, when I did a linear regression model test to further explore the relationship between diabetes coutcome and bmi, I get a clearer story (summarize below)
```

The t-values are -3.159 for the intercept and 8.472 for the bmi coefficient. The t-value for the intercept measures the significantly the intercept deviates from 0 (since the intercept represents the predicted value of the outcome variable when all predictor variables are zero). The -3.159 indicates that the intercept is significantly different from 0, but this value may not have very much meaning for interpreting the data.

However, the t-value for the bmi coefficient (which represents the change in the diabetes outcome for every unit change in bmi) measures the statistical significance of the relationship between bmi and the diabetes outcome. With the t-value for the bmi coefficient as 8.472, it suggests a corresponding low p-value, indicating that there may be a good statistical relationship between bmi and the diabetes outcome. To decide, we must look at the p-values as high t-values corresponding to low p-values indicate statistical significance.

Based on the significance codes, the p-values indicate a very high level of statistical signficance.

More specifically, for the intercept (value of the outcome when the predictor is 0), \*\* indicates high statistical significance, and a p-value of .00165 falls way under my chosen statistical significance level of 0.05, informing me to reject the H0 (null) hypothesis.

For the bmi predictor, the p-value, as indicated by \*\*\*, was of extremely high statistical significance being 2 x 10[-16] (raised to the power of-16) - a very, very small number - far, far below 0.05. And so, this gives me the confidence to reject the H0 hypothesis in favor of the HA hypothesis: There is a statistically significant relationship between the bmi and the diabetes outcome.

From the test results, I can also formulate the linear regression model for predicting the diabetes outcome based on the bmi using the following predictive formula:

  y = -0.21752 + 0.01771x, where y is the predicted diabetes outcome and x is the bmi of the person

  (-0.21752 is the intercept coefficient and the 0.01771 is the bmi coefficient)

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```
